---
output:
  word_document: default
  html_document: default
---

# MIS503 – Final Project
## Warrene, Audrey
### Zillow Home Value Index Analysis

Exercise 1 ----

### Wake County Home Sales

a. What have been the overall trends in Wake County Home Values? -- Increasing, substantially in 2022-2023
b. There were dips in home values in the past 10 years. What years did these occur? -- The most obvious dip in prices is from 2022-2023, which we can see on both the bar chart and the scatterplot. Otherwise, there is not an obvious drop in prices on the charts.
c. Based on the analysis, where would be the least expensive area to purchase home? Most expensive
area? -- Least Expensive: Zebulon. Most Expensive: Cary.
d. What has happened to the overall property values in Apex and Cary in 2023? -- It has slightly gone down since 2022, but is still consistently staying high on the chart. Cary is still more expensive than Apex as well. 

```{r Imports}

library(tidyverse)

SingleFamilyResidenceRental <- read_csv("SingleFamilyResidenceRental.csv")
SingleFamilyResidenceSales <- read_csv("SingleFamilyResidenceSales.csv")

```

```{r}

WakeCountySales <- select(SingleFamilyResidenceSales, RegionName, State, CountyName, Metro, "5/31/2013", "5/31/2014", "5/31/2015", "5/31/2016", "5/31/2017", "5/31/2018", "5/31/2019", "5/31/2020", "5/31/2021", "5/31/2022", "5/31/2023")
WakeCountySales <- filter(WakeCountySales, State=="NC", CountyName=="Wake County")

WakeCountySales

WakeCountySales <- rename(WakeCountySales, "2013" = "5/31/2013")
WakeCountySales <- rename(WakeCountySales, "2014" = "5/31/2014")
WakeCountySales <- rename(WakeCountySales, "2015" = "5/31/2015")
WakeCountySales <- rename(WakeCountySales, "2016" = "5/31/2016")
WakeCountySales <- rename(WakeCountySales, "2017" = "5/31/2017")
WakeCountySales <- rename(WakeCountySales, "2018" = "5/31/2018")
WakeCountySales <- rename(WakeCountySales, "2019" = "5/31/2019")
WakeCountySales <- rename(WakeCountySales, "2020" = "5/31/2020")
WakeCountySales <- rename(WakeCountySales, "2021" = "5/31/2021")
WakeCountySales <- rename(WakeCountySales, "2022" = "5/31/2022")
WakeCountySales <- rename(WakeCountySales, "2023" = "5/31/2023")

WakeCountySales

WakeCountySales <- WakeCountySales %>% pivot_longer(c('2013','2014', '2015', '2016', '2017', '2018', '2019', '2020', '2021', '2022', '2023'),names_to='YR',values_to='ZHVI')

WakeCountySales

```

```{r}

ggplot(WakeCountySales, aes(x = YR, y = ZHVI, color = RegionName)) +
  geom_point() +
  labs(title = "Wake County Home Sales",
       x = "Year",
       y = "Home Values (in $)") +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5)) +
  scale_y_continuous(name="Home Values (in $)", labels = scales::comma)

ggplot(WakeCountySales, aes(x = YR, y = ZHVI, color = RegionName)) +
  geom_col() +
  labs(title = "Wake County Home Sales",
       x = "Year",
       y = "Home Values (in $)") +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5)) +
  scale_y_continuous(name="Home Values (in $)", labels = scales::comma)

```

Exercise 2 ----

### NC Rental Market

a. What has been the overall trend in the rental market around the state? Are there any cities that
have not followed this trend? -- The overall trend is an increase in rental price over the years, especially in 2022 and 2023. All cities have followed this trend, but Fayetteville is following it at a much smaller pace.
b. Where is the most expensive city to rent in? Least expensive? -- Most Expensive: Asheville. Least Expensive: Fayetteville.
c. You are trying to decide between Wilmington and Asheville. Which market has the lowest rent? -- Wilmington has the lower rent between the two. 

```{r}

Rentals <- select(SingleFamilyResidenceRental, RegionName, State, "1/31/2015", "1/31/2016", "1/31/2017", "1/31/2018", "1/31/2019", "1/31/2020", "1/31/2021", "1/31/2022", "1/31/2023")
Rentals <- filter(Rentals, State=="NC", RegionName %in% c("Asheville", "Charlotte", "Durham", "Fayetteville", "Raleigh", "Wilmington"))

Rentals

Rentals <- rename(Rentals, "2015" = "1/31/2015")
Rentals <- rename(Rentals, "2016" = "1/31/2016")
Rentals <- rename(Rentals, "2017" = "1/31/2017")
Rentals <- rename(Rentals, "2018" = "1/31/2018")
Rentals <- rename(Rentals, "2019" = "1/31/2019")
Rentals <- rename(Rentals, "2020" = "1/31/2020")
Rentals <- rename(Rentals, "2021" = "1/31/2021")
Rentals <- rename(Rentals, "2022" = "1/31/2022")
Rentals <- rename(Rentals, "2023" = "1/31/2023")

Rentals

Rentals <- Rentals %>% pivot_longer(c('2015', '2016', '2017', '2018', '2019', '2020', '2021', '2022', '2023'),names_to='YR',values_to='ZHVI')

Rentals

```

```{r}

ggplot(Rentals, aes(x = YR, y = ZHVI, color = RegionName)) +
  geom_point() +
  labs(title = "Rental Costs in NC Cities",
       x = "Year",
       y = "Average Rental Cost") +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5)) +
  scale_y_continuous(name="Average Rental Cost", labels = scales::comma)

ggplot(Rentals, aes(x = YR, y = ZHVI, color = RegionName)) +
  geom_col() +
  labs(title = "Rental Costs in NC Cities",
       x = "Year",
       y = "Average Rental Cost") +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5)) +
  scale_y_continuous(name="Average Rental Cost", labels = scales::comma)

```

Exercise 3 ----

### Home Values in Select Markets

a. According to the results, which market has the lowest median price (represented as horizontal bar in
box plot)? -- Charlotte-Concord-Gastonia
b. The violin plot will show density meaning the wider the plot is, the more observations occur within
that area. Which market has the most density around the median value of homes? -- Asheville
c. The box plot will also show outliers in the various markets. Which metro area had the largest outlier
(i.e., the highest value home sold in the past 10 years)? -- Wilmington

```{r}

NCHomeSales <- select(SingleFamilyResidenceSales, RegionName, State, Metro, "5/31/2013", "5/31/2014", "5/31/2015", "5/31/2016", "5/31/2017", "5/31/2018", "5/31/2019", "5/31/2020", "5/31/2021", "5/31/2022", "5/31/2023")

NCHomeSales <- filter(NCHomeSales, State=="NC", Metro %in% c("Asheville", "Charlotte-Concord-Gastonia", "Raleigh-Cary", "Wilmington"))

NCHomeSales

NCHomeSales <- rename(NCHomeSales, "2013" = "5/31/2013")
NCHomeSales <- rename(NCHomeSales, "2014" = "5/31/2014")
NCHomeSales <- rename(NCHomeSales, "2015" = "5/31/2015")
NCHomeSales <- rename(NCHomeSales, "2016" = "5/31/2016")
NCHomeSales <- rename(NCHomeSales, "2017" = "5/31/2017")
NCHomeSales <- rename(NCHomeSales, "2018" = "5/31/2018")
NCHomeSales <- rename(NCHomeSales, "2019" = "5/31/2019")
NCHomeSales <- rename(NCHomeSales, "2020" = "5/31/2020")
NCHomeSales <- rename(NCHomeSales, "2021" = "5/31/2021")
NCHomeSales <- rename(NCHomeSales, "2022" = "5/31/2022")
NCHomeSales <- rename(NCHomeSales, "2023" = "5/31/2023")

NCHomeSales

NCHomeSales <- NCHomeSales %>% pivot_longer(c('2013','2014', '2015', '2016', '2017', '2018', '2019', '2020', '2021', '2022', '2023'),names_to='YR',values_to='ZHVI')

NCHomeSales <- arrange(NCHomeSales, Metro)

NCHomeSales

```

```{r}

ggplot(NCHomeSales, aes(x = Metro, y = ZHVI)) +
  geom_violin() +
  geom_boxplot(width = 0.1) +
  labs(x = "Metro", 
       y = "Home Values (in $)")

```

Exercise 4 ----

###  Relocation Home Value Comparison

a. Based on your analysis, which city’s housing is most affordable? Least affordable? -- Most Affordable: Houston. Least Affordable: New York.
b. Which cities saw the largest change in prices over the past 5 years? Which city has remained more
consistent (i.e., no huge swings up or down in home values)? -- The city with the largest changes in prices is Denver. The city that has remained more consistent is Chicago, because it has consistently remained at 500,000 dollars.
c. Which cities saw a decline in value during 2023 and which cities remained consistent? -- Decline: New York, Denver, Technically Chicago. The cities remaining most consistent is Houston and technically Chicago, because even though the value did drop, it BARELY dropped. 

```{r}

NationalHomeSales <- select(SingleFamilyResidenceSales, RegionName, State, CountyName, Metro, "5/31/2013", "5/31/2014", "5/31/2015", "5/31/2016", "5/31/2017", "5/31/2018", "5/31/2019", "5/31/2020", "5/31/2021", "5/31/2022", "5/31/2023")

cities <- c("Chicago", "Denver", "Houston", "New York")
counties <- c("Cook County", "Denver County", "Harris County", "Queens County")

NationalHomeSales <- filter(NationalHomeSales, RegionName %in% cities, CountyName %in% counties)

NationalHomeSales

NationalHomeSales <- rename(NationalHomeSales, "2013" = "5/31/2013")
NationalHomeSales <- rename(NationalHomeSales, "2014" = "5/31/2014")
NationalHomeSales <- rename(NationalHomeSales, "2015" = "5/31/2015")
NationalHomeSales <- rename(NationalHomeSales, "2016" = "5/31/2016")
NationalHomeSales <- rename(NationalHomeSales, "2017" = "5/31/2017")
NationalHomeSales <- rename(NationalHomeSales, "2018" = "5/31/2018")
NationalHomeSales <- rename(NationalHomeSales, "2019" = "5/31/2019")
NationalHomeSales <- rename(NationalHomeSales, "2020" = "5/31/2020")
NationalHomeSales <- rename(NationalHomeSales, "2021" = "5/31/2021")
NationalHomeSales <- rename(NationalHomeSales, "2022" = "5/31/2022")
NationalHomeSales <- rename(NationalHomeSales, "2023" = "5/31/2023")

NationalHomeSales

NationalHomeSales <- NationalHomeSales %>% pivot_longer(c('2013','2014', '2015', '2016', '2017', '2018', '2019', '2020', '2021', '2022', '2023'),names_to='YR',values_to='ZHVI')

NationalHomeSales

```

```{r}

ggplot(NationalHomeSales, aes(x = YR, y = ZHVI)) +
  geom_point() +
  labs(x = "Year",
       y = "Home Values (in $)") +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5)) +
  scale_y_continuous(name="Home Values (in $)", labels = scales::comma) +
  facet_wrap(~Metro)

```

Exercise 5 ----

### Future Home Values

a. Which is the only city that is projected to have a decrease in home values in the next 3 months? -- None of them really dropped in home value, but Hoston, TX maybe dropped? (or just remains the same)
b. If you are only concerned about the largest home value increase (by percentage) in the next 12
months, which city would you choose to relocate to? -- Chicaco, IL has the best percentage after 12 months.

```{r Import}

Projections <- read_csv("Projections.csv")

```

```{r}

FutureHomeValues <- select(Projections, RegionName, "2024-04-30", "2024-06-30", "2025-03-31")

FutureHomeValues

cities <- c("Chicago, IL", "Denver, CO", "Houston, TX", "New York, NY")

FutureHomeValues <- filter(FutureHomeValues, RegionName %in% cities)

FutureHomeValues

FutureHomeValues <- rename(FutureHomeValues, "Current" = "2024-04-30")
FutureHomeValues <- rename(FutureHomeValues, "ThreeMonths" = "2024-06-30")
FutureHomeValues <- rename(FutureHomeValues, "TwelveMonths" = "2025-03-31")

FutureHomeValues

FutureHomeValues <- FutureHomeValues %>% pivot_longer(c('Current', 'ThreeMonths', 'TwelveMonths'),names_to='Time',values_to='PercentageChange')

FutureHomeValues

```

```{r}

ggplot(FutureHomeValues, aes(x = Time, y = PercentageChange, color = RegionName)) +
  geom_point()

```

